### 2019-02-11 symmetry_properties.tex
ssr

\subsection{Overview}
\label{sse:res_overview}
Our first result (Theorem~\ref{thm:ncf_symmetric_variables})
provides a characterization of pairs of variables of an NCF
that are part of the same symmetry group. 
This characterization allows us to give a simple closed form
expression (Theorem~\ref{thm:ncf_r_symmetric}) 
for the symmetry level of an NCF specified using
its normalized representation.
Our next result (Theorem~\ref{thm:ncf_strong_asymmetry})
shows that for any NCF $f$ with $n$ variables,
the two statements ``$f$ is strongly asymmetric"
``the symmetry level of $f$ is $n$" are equivalent. 
Our final result (Proposition~\ref{pro:ncf_symmetric})
identifies all the functions that are symmetric and
canalyzing as well as those that are symmetric and NCFs.

@@

\begin{theorem}\label{thm:ncf_symmetric_variables}
Two variables of an NCF $f$ are symmetric iff
in the normalized representation of $f$,
they occur in the same layer and have the same canalyzing value.
\end{theorem}
\noindent
\textbf{Proof:}~
Suppose that two variables of an NCF $f$ occur in the same layer of a
normalized representation, and the lines containing these variables
have the same canalyzing value.  Then in any input assignment, the
values of these two variables can be interchanged, and the evaluation
of the lines in that layer will have the same effect.  Thus, the
variables are symmetric.

For the converse, consider the normalized NCF representation of $f$. 
For purposes of notational simplicity, we assume without loss of generality that the
canalyzing variable in line $i$ of $f$ is $x_i$, $1 \leq i \leq n$.
Let  $x_j$ and $x_k$ be two variables that occur in different
layers of $f$, or occur in the same layer, but with different
canalyzing values.  Without loss of generality, 
assume that $j < k$, so that the line for
$x_j$ occurs above the line for $x_k$.
Suppose that line $i$ of $f$, $1 \leq i \leq n$, is

@@

Now suppose that line $k$ is the only line below line $j$ for which
the canalyzed value is $\overline{b_j}$.  Since line $j$ has
canalyzed value $b_j$, line $k$ is the only line in its layer.
Since $f$ is normalized, the last layer of $f$ contains at least
two lines.  Thus, there is a last layer following the layer containing
line $k$.  The canalyzed value of all the lines in this last layer
is $b_j$.  Thus, for every line $i$, $1 \leq i \leq n$, the value
of variable $x_i$ in assignment $\alpha'$ does not match the
canalyzing value of line $i$.  Consequently, $f(\alpha') =
\overline{b_j}$, the complement of the canalyzed value in the last
layer of $f$.  \QED

\begin{theorem}\label{thm:ncf_r_symmetric}
Suppose the normalized representation of a given NCF contains 
$r_1$ layers with only one distinct canalyzing value,
and $r_2$ layers with two distinct canalyzing values.
Then the function is properly $(r_1 + 2 r_2)$-symmetric.
\end{theorem}

@@

\begin{corollary}\label{cor:ncf_r_symmetric_layers}
(i) An NCF with a normalized nested canalyzing representation consisting
of $q$ layers is $2q$-symmetric, and is not $(q-1)$-symmetric.
(ii) An $r$-symmetric NCF has a normalized nested canalyzing representation
with at most $r-1$ layers, and has an unnormalized nested canalyzing
representation with at most $r$ layers.  
\end{corollary}

@@

For the converse, consider an $n$ variable NCF that is properly $n$-symmetric.
Let $f$ be the normalized NCF representation for the function.  For
purposes of notational simplicity, without loss of generality,
we assume that the canalyzing
variable in line $i$ of $f$ is $x_i$, $1 \leq i \leq n$.  Let $\pi$
be any permutation of $\{1, 2, \ldots, n\}$ \emph{except} the
identity permutation.  We will construct an assignment $(c_1, c_2,
\ldots,  c_n)$ to the variables of $f$ such that $f(c_1, c_2, \ldots,
c_n)$ $\neq$ $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)})$.

@@

Now suppose there is no line below line $k$ for which the canalyzed
value is $\overline{b_k}$.  Then line $k$ occurs in the last layer
of $f$.  Since $f$ is both normalized and properly $n$-symmetric,
this last layer contains exactly two lines.  Thus $k = n-1$ and $k'
= n$.  Note that for all $i$, $c_{\pi(i)}$ does not match the
canalyzing value of line $i$.  Thus, $f(c_{\pi(1)}, c_{\pi(2)},
\ldots, c_{\pi(n)}) = \overline{b_k}$.

@@

Now suppose that line $k'$ is the only line below line $k$ for which
the canalyzed value is $\overline{b_k}$.  Since line $k$ has
canalyzed value $b_k$, line $k'$ is the only line in its layer.
Since $f$ is normalized, there is a last layer following the layer
containing line $k'$.  Thus, for all $i$, $c_{\pi(i)}$ does not
match the canalyzing value of line $i$, so $f(c_{\pi(1)}, c_{\pi(2)},
\ldots, c_{\pi(n)}) = \overline{b_k}$.  \QED

@@

\noindent
\textbf{Proof:}~
Consider a default normalized representation of a strongly asymmetric NCF $f$.
The representation contains $n-1$ layers.
The first $n-2$ layers each contain one line, and the last layer contains two lines.

>>>>

\subsection{Overview}
\label{sse:res_overview}
Our first result (Theorem~\ref{thm:ncf_symmetric_variables})
provides a characterization of pairs of variables of an NCF
that are part of the same symmetry group. 
This characterization allows us to give a simple closed form
expression (Theorem~\ref{thm:ncf_r_symmetric}) 
for the symmetry level of an NCF specified using
its default-normalized representation.
%% We also list some corollaries that follow from the
%% characterization.
Next, we show that for any NCF $f$ with $n$ variables,
the two statements ``$f$ is strongly asymmetric" ~and~
``the symmetry level of $f$ is $n$"~ are equivalent 
(Theorem~\ref{thm:ncf_strong_asymmetry}).
We use this result to derive a closed form expression 
for the number of $n$-variable NCFs that are 
also strongly asymmetric (Theorem~\ref{thm:count_strongly_asymmetric}).
Our final result (Proposition~\ref{pro:ncf_symmetric})
identifies all the functions that are symmetric and
canalyzing as well as those that are symmetric and NCFs.

@@

\begin{theorem}\label{thm:ncf_symmetric_variables}
Two variables of an NCF $f$ are symmetric iff
in the default-normalized representation of $f$,
they occur in the same layer and have the same canalyzing value.
\end{theorem}
\noindent
\textbf{Proof:}~
Suppose that two variables of an NCF $f$ occur in the same layer of a
default-normalized representation, and the lines containing these variables
have the same canalyzing value.  Then in any input assignment, the
values of these two variables can be interchanged, and the evaluation
of the lines in that layer will have the same effect.  Thus, the
variables are symmetric.

For the converse, consider the default-normalized NCF representation of $f$. 
For purposes of notational simplicity, we assume without loss of generality that the
canalyzing variable in line $i$ of $f$ is $x_i$, $1 \leq i \leq n$.
Let  $x_j$ and $x_k$ be two variables that occur in different
layers of $f$, or occur in the same layer, but with different
canalyzing values.  Without loss of generality, 
assume that $j < k$, so that the line for
$x_j$ occurs above the line for $x_k$.
Suppose that line $i$ of $f$, $1 \leq i \leq n$, is

@@

Now suppose that line $k$ is the only line below line $j$ for which
the canalyzed value is $\overline{b_j}$.  Since line $j$ has
canalyzed value $b_j$, line $k$ is the only line in its layer.
Since $f$ is default-normalized, the last layer of $f$ contains at least
two lines.  Thus, there is a last layer following the layer containing
line $k$.  The canalyzed value of all the lines in this last layer
is $b_j$.  Thus, for every line $i$, $1 \leq i \leq n$, the value
of variable $x_i$ in assignment $\alpha'$ does not match the
canalyzing value of line $i$.  Consequently, $f(\alpha') =
\overline{b_j}$, the complement of the canalyzed value in the last
layer of $f$.  \QED

\begin{theorem}\label{thm:ncf_r_symmetric}
Suppose the default-normalized representation of a given NCF contains 
$r_1$ layers with only one distinct canalyzing value,
and $r_2$ layers with two distinct canalyzing values.
Then the function is properly $(r_1 + 2 r_2)$-symmetric.
\end{theorem}

@@

\begin{corollary}\label{cor:ncf_r_symmetric_layers}
(i) An NCF with a default-normalized nested canalyzing representation consisting
of $q$ layers is $2q$-symmetric, and is not $(q-1)$-symmetric.
(ii) An $r$-symmetric NCF has a default-normalized nested canalyzing representation
with at most $r-1$ layers, and has an unnormalized nested canalyzing
representation with at most $r$ layers.  
\end{corollary}

@@

For the converse, consider an $n$ variable NCF that is properly $n$-symmetric.
Let $f$ be the default-normalized NCF representation for the function.  For
purposes of notational simplicity, without loss of generality,
we assume that the canalyzing
variable in line $i$ of $f$ is $x_i$, $1 \leq i \leq n$.  Let $\pi$
be any permutation of $\{1, 2, \ldots, n\}$ \emph{except} the
identity permutation.  We will construct an assignment $(c_1, c_2,
\ldots,  c_n)$ to the variables of $f$ such that $f(c_1, c_2, \ldots,
c_n)$ $\neq$ $f(c_{\pi(1)}, c_{\pi(2)}, \ldots, c_{\pi(n)})$.

@@

Now suppose there is no line below line $k$ for which the canalyzed
value is $\overline{b_k}$.  Then line $k$ occurs in the last layer
of $f$.  Since $f$ is both default-normalized and properly $n$-symmetric,
this last layer contains exactly two lines.  Thus $k = n-1$ and $k'
= n$.  Note that for all $i$, $c_{\pi(i)}$ does not match the
canalyzing value of line $i$.  Thus, $f(c_{\pi(1)}, c_{\pi(2)},
\ldots, c_{\pi(n)}) = \overline{b_k}$.

@@

Now suppose that line $k'$ is the only line below line $k$ for which
the canalyzed value is $\overline{b_k}$.  Since line $k$ has
canalyzed value $b_k$, line $k'$ is the only line in its layer.
Since $f$ is default-normalized, there is a last layer following the layer
containing line $k'$.  Thus, for all $i$, $c_{\pi(i)}$ does not
match the canalyzing value of line $i$, so $f(c_{\pi(1)}, c_{\pi(2)},
\ldots, c_{\pi(n)}) = \overline{b_k}$.  \QED

@@

\noindent
\textbf{Proof:}~
Consider a default-normalized representation of a strongly asymmetric NCF $f$.
The representation contains $n-1$ layers.
The first $n-2$ layers each contain one line, and the last layer contains two lines.
